% !TEX root = paper.tex

This appendix contains some additional structural results about Burkholder functions which may be useful for algorithm designers.

\begin{proposition}
\label{prop:algebra}
The following statements are true:
\begin{enumerate}
\item Given a \Bfun function $\burk$, if we define the $X_t = \burk(\sum_{j=1}^t \suff(z_j,\delta_j))$, then for any real-valued martingale difference sequence $\delta_t$s and predictable $z_t$s, $(X_t)_{t \ge 0}$ is a supermartingale with $\mathbb{E}[X_0] \le 0$.
\item Any convex combination of \Bfun functions is a \Bfun function.
\item The minimum of a family of \Bfun functions is a \Bfun function.
\item Suppose we have a finite set $A$ that indexes a family of functions $V_a:\cT\to\bbR$, each of which belongs to a sufficient statistic pair $(\suff, V_a)$ for some regret inequality of interest, and suppose each $V_a$ has a corresponding \Bfun function $\burk_a$.  Then the following probabilistic inequality is true:
$$
\mathbb{E}\left[\max_{a \in A} \left\{V_a\prn*{\sum_{t=1}^n \suff(z_t,\delta_t)} - \eta n C[a] \right\}\right]  \le \frac{1}{\eta}\log |A|,
$$
where $C[a] = \sup_{\tau, z ,\alpha} (\burk_a(\tau + \suff(z,\alpha)) - \burk_a(\tau))^2$. Note that $C \in \mathbb{R}^A$ may be thought as a sufficient statistic, though it is fixed and does not depend on instances.
Furthermore, a \Bfun function $\burk: \T \times \mathbb{R}^A \to \reals$ that certifies this inequality is:
\begin{equation}
\label{eq:burkholder_meta}
\burk(\tau, \gamma) = \frac{1}{\eta} \log\left(\sum_{a \in A} \exp\left(\eta \burk_a(\tau) - \eta^2 \gamma[a] \right)\right)  - \frac{\log|A|}{\eta}
\end{equation}
\end{enumerate}
\end{proposition}
\begin{proof}[\pfref{prop:algebra}]
The first statement follows from property \propthree{} of the Burkholder function $\burk$, which immediately implies that it is a supermartingale. The second statement is trivial. To prove the third statement it suffices to verify property \propthree, which holds due to concavity of the minimum. 

We now prove the fourth statement. Given a family of \Bfun functions $\crl*{\burk_a}_{a\in{}A}$, define a new \Bfun function $\burk: \T \times \mathbb{R}^A \to\reals$ as:
$$
\burk(\tau, \gamma) = \frac{1}{\eta} \log\left(\sum_{a \in A} \exp\left(\eta \burk_a(\tau) - \eta^2 \gamma[a] \right)\right)  - \frac{\log|A|}{\eta}.
$$
whose sufficient statistics are the original sufficient statistic of the family of $V_a$s along with an additional $|A|$-dimensional real vector, for which one coordinate per $a \in A$ will be used to represent $C[a] = \sup_{\tau, z ,\alpha} (\burk_a(\tau + \suff(z,\alpha)) - \burk_a(\tau))^2$ (note that this is a vacuous statistic as it is constant for each instance). Property \propthree{} for $\burk$ holds as follows:
\begin{align*}
 \En_\alpha &\burk\left((\tau,\gamma) + (\suff(z,\alpha) , C) \right) \\
 & = \frac{1}{\eta} \En_\alpha \log\left(\sum_{a \in A} \exp\left(\eta \burk_a(\tau + \suff(z,\alpha)) - \eta^2 \gamma[a] - \eta^2 C[a]\right)\right) - \frac{\log|A|}{\eta} \\
 & \le  \frac{1}{\eta}  \log\left(\sum_{a \in A} \En_\alpha \exp\left(\eta \burk_a(\tau + \suff(z,\alpha))  - \eta^2 \gamma[a] - \eta^2 C[a] \right)\right) - \frac{\log|A|}{\eta} \\
 & =  \frac{1}{\eta}  \log\left(\sum_{a \in A} \En_\alpha \exp\left(\eta \left(\burk_a(\tau + \suff(z,\alpha)) - \burk_a(\tau) \right) + \eta\burk_a(\tau)  - \eta^2 \gamma[a] - \eta^2 C[a]\right)\right) - \frac{\log|A|}{\eta}.
\end{align*}
Now note that by property \propthree{} of the \Bfun functions $\crl*{\burk_a}_{a\in{}A}$, the random variable $X_a = \left(\burk_a(\tau + \suff(z,\alpha)) - \burk_a(\tau) \right)$ is such that $\En_{\alpha}[X_a] \le 0$. Further from our assumption we have that $|X_a|^2 \le C[a]$. Hence, the standard mgf bound implies $\mathbb{E}_{\alpha}[\exp(\eta X_a)] \le \exp(\eta^2 C[a]/2)$.
\begin{align*}
 & \le  \frac{1}{\eta}  \log\left(\sum_{a \in A} \exp\left(\eta\burk_a(\tau)  + \frac{\eta^2}{2} C[a]  - \eta^2 \gamma[a] - \eta^2 C[a] \right)\right) - \frac{\log|A|}{\eta} \\
 & \le \frac{1}{\eta}  \log\left(\sum_{a \in A} \exp\left(\eta\burk_a(\tau)   - \eta^2 \gamma[a] \right)\right) - \frac{\log|A|}{\eta}.
\end{align*}
For property \propone{} it can be seen immediately that $\burk(0) \le 0$. Property \proptwo{} holds via
\begin{align*}
 \burk(\tau,\gamma) &= \frac{1}{\eta} \log\left(\sum_{a \in A} \exp\left(\eta \burk_a(\tau) - \eta^2 \gamma[a] \right)\right)  - \frac{\log|A|}{\eta}\\
& \ge \max_{a \in A}\left\{ \burk_a(\tau) - \eta \gamma[a]\right\} - \frac{\log|A|}{\eta} ~~~~~~ \textrm{(softmax upper bounds max)}\\
& \ge \max_{a \in A}\left\{ V_a(\tau) - \eta \gamma[a]\right\} - \frac{\log|A|}{\eta}.
\end{align*}
\end{proof}
We remark that one uses non-additive sufficient statistics as discussed in \pref{app:discussion}, then one can make the bound implied by the Burkholder function $\burk$ above more data-dependent by replacing $C[a]$ with $\sup_{\delta} \left(\burk_a(\tau + \suff(z,\delta)) - \burk_a(\tau) \right)^2$ for each $a$. 
