
\section{Conclusions}
%
We have investigated the setting of $d$-delayed composite anonymous feedback as applied to nonstochastic bandits. Composite anonymous feedback lends itself to formalize scenarios where the actions perfomed by the online decision maker produce long-lasting effects that combine additively over time. A general reduction technique was introduced that enables the convertion of a (backward stable) algorithm working in a standard bandit framework into one working in the composite feedback framework. 
In the case of $K$-armed bandits, we relied on a lower bound for bandit linear optimization in the probability simplex to show that no algorithm in the composite feedback framework can do better than $\scO(\sqrt{dKT})$. In turn, up to log factors, this is what we obtain as an upper bound by applying our reduction to the standard Exp3 algorithm.

We showed the generality and flexibility of our conversion technique by further applying it to Combinatorial Bandits (the Exp2 algorithm) and to Bandit Convex Optimization (the self-concordant barrier-based algorithms by \cite{ahr12} and \cite{st11}) with smooth/linear loss functions. In the latter case, the resulting Bandit Convex Optimization algorithm for composite anonymous feedback turned out to achieve regret bounds with higher rates than those of the corresponding base bandit algorithms, and we suspect that there is margin for improvement in our stability analysis of Lemma \ref{l:stabilityconvexscrible}.


\iffalse
************************
We then extended our results to the Bandit Convex Optimization framework, thereby showing the generality and flexibility of our conversion technique. As an example, we applied 

We Our reduction is flexible

 that works under the assumption of an oblivious adversary generating loss components. 

provided a general reduction that does such and such

in the case of k armed upper and lower bound, matching up to log factors. lower bound reduces to bandit linear optimization




flexibility, extension to combinatorial and bco, an example of application with beta smooth and self-concordant barriers.

Our analysis of bco is almost certainly suboptimal,
************************
\fi


