\documentclass[final,12pt]{colt2018} % Anonymized submission
% \documentclass{colt2017} % Include author names
\input{preamble}
% The following packages will be automatically loaded:
% amsmath, amssymb, natbib, graphicx, url, algorithm2e

\title[Actively Avoiding Nonsense in Generative Models]{Actively Avoiding Nonsense in Generative Models}
\usepackage{times}
\usepackage{algorithmic}
 % Use \Name{Author Name} to specify the name.
 % If the surname contains spaces, enclose the surname
 % in braces, e.g. \Name{John {Smith Jones}} similarly
 % if the name has a "von" part, e.g \Name{Jane {de Winter}}.
 % If the first letter in the forenames is a diacritic
 % enclose the diacritic in braces, e.g. \Name{{\'E}louise Smith}

 % Two authors with the same address
  % \coltauthor{\Name{Author Name1} \Email{abc@sample.com}\and
  %  \Name{Author Name2} \Email{xyz@sample.com}\\
  %  \addr Address}

 % Three or more authors with the same address:
 % \coltauthor{\Name{Author Name1} \Email{an1@sample.com}\\
 %  \Name{Author Name2} \Email{an2@sample.com}\\
 %  \Name{Author Name3} \Email{an3@sample.com}\\
 %  \addr Address}


 % Authors with different addresses:
 \coltauthor{\Name{Steve Hanneke} \Email{steve.hanneke@gmail.com}\\
 \addr Princeton, NJ
 \AND
 \Name{Adam Tauman Kalai} \Email{adum@microsoft.com}\\%I'm happy to be last author if you choose not to go alphabetically.
 \addr Microsoft Research, New England
 \AND
 \Name{Gautam Kamath} \Email{g@csail.mit.edu}\\
 \addr EECS \& CSAIL, MIT
 \AND
 \Name{Christos Tzamos} \Email{chtzamos@microsoft.com}\\
 \addr Microsoft Research, New England
 }

\begin{document}

\maketitle

\begin{abstract}
A generative model may generate utter nonsense when it is fit to maximize the likelihood of observed data. This happens due to ``model error,'' i.e., when the true data generating distribution does not fit within the class of generative models being learned. To address this, we propose a model of active distribution learning using a binary invalidity oracle that identifies some examples as clearly invalid, together with random positive examples sampled from the true distribution. The goal is to maximize the likelihood of the positive examples subject to the constraint of (almost) never generating examples labeled invalid by the oracle. Guarantees are agnostic compared to a class of probability distributions. We first show that proper learning may require exponentially many queries to the invalidity oracle. We then give an improper distribution learning algorithm that uses only polynomially many queries. 
%Our algorithm is a polynomial-time reduction to a supervised generative modeling problem.
\end{abstract}

\begin{keywords}
Generative models, active learning, statistical learning
\end{keywords}

\input{intro}
\input{preliminaries}
\input{proper}
\input{improper}
\input{extensions}

\bibliography{biblio}

\appendix
\input{proper-lb-proof}
\input{partial-proofs}
\input{vc-classes}

\end{document}
