
\section{Summary}

Minimizing a convex function over a convex set is a fundamental problem
with many applications. The problem stands at the forefront of polynomial-time
tractability and its study has lead to the development of numerous
general algorithmic techniques. In recent years, improvements to important
special cases (e.g., maxflow) have been closely related to ideas and
improvements for the general problem \citep{christiano2011electrical,sherman2013nearly,madry2013navigating,kelner2014almost,lee2013new,lee2014path,lee2015efficient,lee2015faster,DBLP:conf/stoc/Sherman17}. 

Here we consider the very general setting where the objective function
and feasible region are both presented only as oracles that can be
queried, specifically an \emph{evaluation oracle} for the function
and a \emph{membership oracle} for the set. We study the problem of
minimizing a convex function over a convex set provided only these
oracles as well as bounds $0<r<R$ and a point $x_{0}\in K$ s.t.
$B(x_{0},r)\subseteq K\subseteq B(x_{0},R)$ where $B(x_{0},r)$ is
the ball of radius $r$ centered at $x_{0}\in\R^{n}$. 

It is well-known that with a stronger \emph{separation} oracle for
the set (and subgradient oracle for the function), this problem can
be solved with $\tilde{O}(n)$ oracle queries using any of \cite{Va96,BV04,lee2015faster}
or with $\tilde{O}(n^{2})$ queries by the classic ellipsoid algorithm
\cite{GLS}. Moreover, it is known that the problem can be solved
with only evaluation and membership oracles through reductions shown
by Gr\"{o}tschel, Lovasz and Schrijver in their classic book \cite{GLS}.
However, the reduction in \cite{GLS} appears to take at least $n^{10}$
calls to the membership oracle. This has been improved using the random
walk method and simulated annealing to $n^{4.5}$ \cite{KV06,LV06}
and  \cite{AbernethyH16}  provides further improvements of up to
a factor of $\sqrt{n}$ for more structured convex sets. 

Our main result in this paper is an algorithm that minimizes a convex
function over a convex set using only $\tilde{O}(n^{2})$ membership
and evaluation queries. Interestingly, we obtain this result by first
showing that we can implement a separation oracle for a convex set
and a subgradient oracle for a function using only $\tilde{O}(n)$
membership queries and then
using the known reduction from optimization to separation.
We state the result informally below. 

\begin{thm}
Let $K$ be a convex set specified by a membership oracle, a point
$x_{0}\in\R^{n}$, and numbers $0<r<R$ such that $B(x_{0},r)\subseteq K\subseteq B(x_{0},R)$.
For any convex function $f$ given by an evaluation oracle and
$\epsilon>0$, there is a randomized algorithm that computes a point
$z\in B(K,\epsilon)$ such that 
\[
f(z)\le\min_{x\in K}f(x)+\epsilon\left(\max_{x\in K}f(x)-\min_{x\in K}f(x)\right)
\]
with constant probability using $O(n^{2}\log^{O(1)}(\frac{nR}{\epsilon r}))$
calls to the membership oracle and evaluation oracle and $O(n^{3}\log^{O(1)}(\frac{nR}{\epsilon r}))$
total arithmetic operations. 
\end{thm}

See the full version for the formal statement, proofs, comparison to prior work, and further consequences of this result.

